---
title: "Worksheet 3 : Measures of Central Tendency"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(readr)
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Wrap-Up Variables
- Measures of Central Tendency

---
  
## Wrap-Up Variables

## Measures of Central Tendency



---

## Class Activity

3. Go to https://www.statisticshowto.com/probability-and-statistics/descriptive-statistics/misleading-graphs/ and use the graphs as inspiration to determine some bad practices when displaying data.

<div>
:::{#answer}

ANSWER HERE

:::
</div>

4. How many variables are there in the `surveySP24` data set? How many observations? What is the largest number of grapes someone thinks they can fit in their mouth?

<div>
:::{#answer}

ANSWER HERE

:::
</div>

5. For each variable in `surveySP24`, identify what type it is: numerical discrete, numberical continuous, categorical nominal,  or categorical ordinal?

<div>
:::{#answer}

ANSWER HERE

:::
</div>

6. Install the package `fivethirtyeight` in the console, load it in the code chunk below and save the data biopics locally.
```{r}
#insert code here
```


7. Find the mean of box office draws (`box_office`) for biopics in this dataset
```{r}
#insert code here
```

8. Find the median of box office draws (`box_office`) for biopics in this dataset
```{r}
#insert code here
```

9. Why are the mean and median so different
```{r}
#insert code here
```
