---
title: "Worksheet 17 : t-distribution and proportions"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Sample mean with known $\sigma$
- Sample mean with unknown $\sigma$
- Proportions


---


## Sample mean with known $\sigma$ 

As your sample size $n$ increases, the sampling distribution for the sample mean approaches a normal curve:

$$ \bar{X} \sim N \left(\mu, \frac{\sigma}{\sqrt{n}} \right) $$

### Conditions 

We have been secretly been making assumptions about our data when computing confidence intervals

From now on, before computing a confidence interval using critical $z$-scores we need to check the following conditions:

+ **Independence** 
+ **Normality**

  
#### Example 

A study wants to determine the average amount of sleep students get at Davidson college. A sample of $900$ first year students  were polled on how much sleep they get per night. The average amount of sleep from this sample was found to be $5.7$ hours a night. The population standard deviation is known to be $1.25$ hours. Construct a $95 \%$ confidence interval for the true population mean or explain what conditions prevent you from applying our methods.

```{r}

```


#### Example 

A study wants to determine the average amount of sleep students get at Davidson college. A simple random sample of $20$ students is taken and the average amount of sleep from this sample was found to be $5.7$ hours a night. The population standard deviation is known to be $1.25$ hours. Construct a $95 \%$ confidence interval for the true population mean or explain what conditions prevent you from applying our methods.

```{r}

```

#### Example 

A study wants to determine the average amount of sleep students get at Davidson college. A simple random sample of $20$ students reported the sleep habits displayed below. The population standard deviation is known to be $1.25$ hours. Construct a $95 \%$ confidence interval for the true population mean or explain what conditions prevent you from applying our methods.

| Hours | count |
|:--------: | :-------: |
| 5  | 1    |
| 5.5  | 0    |
| 6  | 1    |
| 6.5  | 2    |
| 7 | 5     |
| 7.5  | 1  |
| 8 | 3 |
| 8.5 | 3 |
| 9 | 4 |

```{r}


```

#### Example 

A study wants to determine the average amount of sleep students get at Davidson college. A simple random sample of $20$ students reported the sleep habits displayed below. The population is assumed to be approximately normal and the population standard deviation is known to be $1.25$ hours. Construct a $95 \%$ confidence interval for the true population mean or explain what conditions prevent you from applying our methods.

| Hours | count |
|:--------: | :-------: |
| 5  | 1    |
| 5.5  | 0    |
| 6  | 1    |
| 6.5  | 2    |
| 7 | 5     |
| 7.5  | 1  |
| 8 | 3 |
| 8.5 | 3 |
| 9 | 4 |

```{r}


```

## Sample mean with unknown $\sigma$ 


### The t-distribution


```{r}
df <- data.frame(x=seq(-5,5,.1))

ggplot(df, aes(x=x)) + 
  stat_function(fun=dt, args = 10, aes(color="df =10")) + 
  stat_function(fun=dt, args = 20,aes(color="df =20")) +
  stat_function(fun=dt, args = 1,aes(color="df = 1")) +
  stat_function(fun=dt, args = 2,aes(color="df = 2")) +
  stat_function(fun=dnorm, args = c(0,1),aes(color="normal")) +
  scale_color_manual("Degrees of Freedom", values = c("red", "blue", "green", "orange","black")) +
  theme_minimal() +
  labs(title = "t-distributions")
  

```


### Conditions

To use the $t$-distribution the same conditions need to be met:

+ Independence
 
+ Normality


## Population Proportion 


#### Example 

Suppose we want to estimate the percentage of adults in a large city that use facebook. Of five hundred randomly sampled adults $421$ say they use facebook. Compute a $90 \%$ confidence interval for the true proportion of adults in the city that use facebook. 

```{r}

```

### Conditions

+ Independence
+ Success-Failure 


1. A study is conducted to analyze internet privacy for teens. A group of $50$ teens is randomly sampled and $7$ report clicking on links from unknown email addresses.  Construct a $90 \%$ confidence interval for the true proportion of teens that click on links from unknown email addresses or explain what conditions prevent you from applying our methods.

```{r}

```

2. A study is conducted to analyze internet privacy for teens. A group of $50$ teens is randomly sampled and $7$ report clicking on links from unknown email addresses.  Construct a $90 \%$ confidence interval for the true proportion of teens that click on links from unknown email addresses or explain what conditions prevent you from applying our methods.

```{r}

```

3. Suppose scores on a statistics exam are normally distributed with an unknown population mean and a population standard deviation of $3$ points. A random sample of $16$ scores is taken and gives a smample mean of $68$. Construct a $95 \%$ confidence interval for the true population mean or explain what conditions prevent you from applying our methods.

```{r}

```

4. Suppose the distribution of average pizza delivery are skewed right with an unknown population mean and  population standard deviation. A random sample of $128$ pizza delivery restaurants is taken and has a sample mean delivery time of $36$ minutes with a standard deviation of $8$ minutes. Construct a $95 \%$ confidence interval for the true population mean or explain what conditions prevent you from applying our methods.

```{r}

```

