---
title: "Worksheet 8 : Probability"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Probability
- Total Probability
- Fundamental Counting Principle

---


## Review - Probability

## Probability

+ Probability is the language of likelihoods

### Random Variables

+ A **random variable** is an object that depends on the outcome of a random event, usually denoted by $X$
+ Examples of a random variable: the face showing for a dice roll, the number of $5$'s in $10$ die rolls, the number of children in a family, the amount of time before a light bulb goes out.
+ there are discrete random variables and continuous random variables.



1. What is the sample space if two different integers are chosen between $1$ and $5$ and listed in increasing order?

<div>
:::{#answer}
answer here
:::
</div>

2. What is the sample space if three dice are rolled and the sum is computed?

<div>
:::{#answer}
answer here
:::
</div>

3. Choose $2$ marbles from a bag with $2$ red marbles, $2$ blue marbles, and $1$ green marble.

<div>
:::{#answer}
answer here
:::
</div>



4. If roll two 4-sided dice, what is the probability of getting doubles?

<div>
:::{#answer}

Answer here

:::
</div>

5. What is the probability of getting at least one head when flipping a coin twice?

<div>
:::{#answer}

Answer here

:::
</div>

## Fundamental Counting Principle



## Probability Distribution


6. Is the following graph a probability distribution describing flipping two fair coins. Why or why not?

```{r}
coin <- data.frame(result = c("2 heads", "1 heads"), probability = c(1/4,3/4))
ggplot(coin, aes(x=result,y=probability))+geom_bar(stat="identity")
```

<div>
:::{#answer}

Answer here

:::
</div>

7. What is the probability of pulling two red marbles from a bag with 2 red marbles, 2 blue marbles, and 1 green marble?

<div>
:::{#answer}

Answer here

:::
</div>

8. On a six sided fair die, what is the probability of rolling an even number or a one?

<div>
:::{#answer}

Answer here

:::
</div>

9. What is the probability of getting at least one heads if you flip a coin three times?

<div>
:::{#answer}

Answer here

:::
</div>

10. From the very first question on this worksheet, what is the probability of the first number being a $2$? What is the probability of the first number being a $5$?

<div>
:::{#answer}

Answer here

:::
</div>