---
title: "Worksheet 9 : Conditional Probability"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Basics
- Pairs of Events
- Disjoint and Independent Events
- Complement


---


## Review - Probability

## Probability

+ Recall that if every outcome is equally likely in a sample space, then 

$$P( X = \text{event}) = \frac{\text{# of successful outcomes}}{\text{total # of outcomes}}$$

+ If you are ever stuck on a probability problem, try listing out all the possible events.
+ Then all you have to do is count the successful ones and divide by the total.
+ A probability distribution is a description of the probabilities of a sample space where:
  - one event must occur
  - the probabilities add to 1
  
1. Describe the probability distribution for counting the number of heads in three coin flips.


<div>
:::{#answer}

Answer here

:::
</div>



## Pairs of events

2. What is the probability of rolling a number greater than three or rolling a four?

<div>
:::{#answer}

Answer here

:::
</div>


## Disjoint and Independent events

Two events are called **disjoint** if they can't both happen at the same time. 

+ heads and tails are disjoint events
+ an even die roll and a die roll that shows a 3 are disjoint
+ drawing a seven from a deck of cards and drawing a heart from a deck of cards are NOT disjoint. You can get the seven of hearts

Two events are called **independent** if the outcome of one does not affect the probability of the other.

+ consecutive coin flips are independent 
+ drawing a two cards from a deck are NOT independent.


3. If events $A$ and $B$ are **disjoint**, $P(A)=.3$ and $P(B)=.4$, what is $P(A \text{ and } B)$?

<div>
:::{#answer}

Answer here

:::
</div>

4. If events $A$ and $B$ are **independent**, $P(A)=.3$ and $P(B)=.4$, what is $P(A \text{ and } B)$?

<div>
:::{#answer}

Answer here

:::
</div>


## Complement 



Given a sample space 

$$S = \{ a, b,c,d,e,f \}$$

and event 

$$A = \{a,d,f \}$$

3. What is $A^c$?

<div>
:::{#answer}

Answer here

:::
</div>

4. What is $P(A) + P(A^c)$?

<div>
:::{#answer}

Answer here

:::
</div>


5. Find the probability that two people in our class have the same birthday.

<div>
:::{#answer}

Answer here

:::
</div>

---

6. Describe the probability distribution of rolling two dice and finding the sum?

<div>
:::{#answer}
answer here
:::
</div>

7. Plot the probability distribution from the last problem.

```{r}
# insert code here
```

8. What is the probability of getting at least two heads when flipping a coin ten times?

<div>
:::{#answer}
answer here
:::
</div>

9. Suppose you flip a fair coin twice, answer the following questions

+ What is the probability of event **A**: getting at most one heads

<div>
:::{#answer}
answer here
:::
</div>

+ What is the probability of event **B**: getting two of the same result (two heads or two tails)

<div>
:::{#answer}
answer here
:::
</div>

+ What is the probability of event **C**: getting heads on the first flip and tails on the second.

<div>
:::{#answer}
answer here
:::
</div>

+ Are events $A$ and $B$ disjoint, independent, both, or neither?

<div>
:::{#answer}
answer here
:::
</div>

+ Are events $A$ and $C$ disjoint, independent, both, or neither?

<div>
:::{#answer}
answer here
:::
</div>

+ Are events $B$ and $C$ disjoint, independent, both, or neither?

<div>
:::{#answer}
answer here
:::
</div>

