---
title: "Lab 7 - Rarity of Events"
author: Student 1 Name, Student 2 Name
output:
  html_document:
    css: ./style.css
    toc: no
    toc_float: no
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(openintro)
library(tidyverse)
set.seed(123)
```

## Instructions

In this lab, you will investigate data and give an explanation of how rare the data is given our expectations. . 

### How to **complete** this assignment.

* Attempt each exercise in order.

* Please follow the Davidson Honor Code and rules from the course syllabus regarding seeking help with this assignment.

* This assignment is <font color="darkred">**due Friday, March 15, no later than 5:00 pm Eastern**</font>. Late submissions are not accepted without prior approval. 

---


Your investigation will focus on the performance of one player: Kobe Bryant of the Los Angeles Lakers. His performance against the Orlando Magic in the 2009 NBA Finals earned him the title Most Valuable Player and many spectators commented on how he appeared to show a hot hand. Having a *hot hand* means that Kobe would tend to go on shooting streaks where he would make lots of shots in a row. You will investigate this belief and make a determination about whether you believe Kobe exhibited the hot hand phenomenon in this game. The data file we’ll use is called `kobe_basket` and is part of the openintro library. 

##### **Exercise 1. Save the `kobe_basket` dataset from the `openintro` package locally. Open the data set from your environment and take a look at it.**

```{r}
#insert code here
```

In this data frame, every row records a shot taken by Kobe Bryant. If he hit the shot (made a basket), a hit, H, is recorded in the column named shot, otherwise a miss, M, is recorded.

Just looking at the string of hits and misses, it can be difficult to gauge whether or not it seems like Kobe was shooting with a hot hand. One way we can approach this is by considering the belief that *hot hand* shooters tend to go on shooting streaks. For this lab, **we define the length of a shooting streak to be the number of consecutive baskets made until a miss occurs**.

For example, in Game 1 Kobe had the following sequence of hits and misses from his nine shot attempts in the first quarter:

$$ H \ M \ | \ M \ | \ H \ H \ M \ | \ M \ | \ M \ | \ M $$

##### **Exercise 2. (A) Insert code that will display the first nine shots in the data frame in order to verify the above sequence. (B) What does the symbol "|" represent in the sequence of hits and misses? **

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>



##### **Exercise 3. (A) What does a streak length of 1 mean, i.e. how many hits and misses are in a streak of 1? What about a streak length of 0? (B) Counting streak lengths manually for all $133$ shots would get tedious, so we’ll use a custom function `calc_streak` to calculate them. Call the function `calc_streak()` and plug in the shot column from the data frame. (C) Save the result from part (B) locally in a vector called `length`. (D) Create a bar graph displaying `length` using `ggplot` and `geom_bar()`. **

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div> 

##### **Exercise 4. (A) Describe the distribution of shot streak lengths. (B) What is Kobe's average streak length, save this locally as `kobe_avg_streak`? What is Kobe's largest streak length?**

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div> 


---

##### Compared to What?

We’ve seen that Kobe had some long shooting streaks, but are they long enough to support the belief that he had a hot hand? What can we compare them to?

To answer these questions, let’s return to the idea of independence. Two processes are independent if the outcome of one process doesn’t affect the outcome of the second. If each shot that a player takes is an independent process, having made or missed your first shot will not affect the probability that you will make or miss your second shot.

A shooter with a hot hand will have shots that are not independent of one another. Specifically, if the shooter makes their first shot, the hot hand model says they will have a higher probability of making their second shot.

Let’s suppose for a moment that the hot hand model is valid for Kobe. During his career, the percentage of time Kobe makes a basket (i.e. his shooting percentage) is about $45 \%$. That is, if $H$ represents a made shot, then in probability notation,


$$P(\text{ H }) = 0.45$$

If he makes the first shot and has a hot hand (i.e. we are assuming the shots are *not* independent), then the probability that he makes his second shot would go up to, let’s say to $60 \%$,

$$P(\text{second shot is }H | \text{first shot is } H) = 0.60$$.

On the other hand, if the second shot is **independent** of the first shot then we would expect

$$P(\text{second shot is }H | \text{first shot is } H) = 0.45$$.

This would mean that the second shot is **unaffected** by the first shot. This is exactly our definition of what it means for two events to be independent. 

##### **Exercise 5. Suppose event that for two events $A$ and $B$ we know that $P(A) = .3$ and $P(B) = .4$. Suppose $A$ and $B$ are independent events and compute $P(A|B)$ and $P(B|A)$**

<div>
:::{#answer}

ANSWER HERE

:::
</div> 


Now that we’ve phrased the situation in terms of independent shots, let’s return to the question: how do we tell if Kobe’s shooting streaks are long enough to indicate that he has a hot hand? We can simulate a basketball player in `R` where we *know* the player *does not* have a hot hand and compare the simulated player to Kobe.

##### More Simulations in R

While we don’t have any data from a shooter we know to have independent shots, that sort of data is very easy to simulate in `R`. We ran some simulations before where we sampled from the diamond data. As a simple reminder example, you can simulate flipping a fair coin with the following:

```{r}
coin_outcomes <- c("heads", "tails")

sample(coin_outcomes, size = 1, replace = TRUE)
```

The vector `coin_outcomes` can be thought of as a hat with two slips of paper in it: one slip says `heads` and the other says `tails`. The function `sample` draws one slip from the hat and tells us if it was a head or a tail (the `replace = TRUE` argument indicates we put the slip of paper back in the hat before drawing again).. 

If you wanted to simulate flipping a fair coin 100 times, you could either run the function 100 times or, more simply, adjust the `size` argument, which governs how many samples to draw Save the resulting vector of heads and tails in a new object called `sim_fair_coin`.

```{r}
sim_fair_coin <- sample(coin_outcomes, size = 100, replace = TRUE)
```

To view the results of this simulation, type the name of the object and then use `table` to count up the number of heads and tails.

```{r}
sim_fair_coin

table(sim_fair_coin)
```

Not all coins are fair and we often encounter situations where the outcomes aren't equally likely. The `sample()` function treats each outcome equally by default, but we can adjust that. Say we’re trying to simulate an unfair coin that we know only lands heads $20 \%$ of the time. We can adjust for this by adding an argument called prob, which provides a vector of two probability weights.

```{r}
sim_unfair_coin <- sample(coin_outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
```

The `prob=c(0.2, 0.8)` indicates that for the two elements in the `outcomes` vector, we want to select the first one, `heads`, with probability $0.2$ and the second one, `tails` with probability $0.8$. Another way of thinking about this is to think of the outcome space as a bag of 10 chips, where 2 chips are labeled “head” and 8 chips “tail”. Therefore at each draw, the probability of drawing a chip that says “head” is $20 \%$, and “tail” is $80 \%$.

---




##### **Exercise 6. Run the simulation of flipping the unfair coin 100 times, how many flips came up heads? How many heads should you expect from this unfair coin? Is the simulation close to the expectation and what could you do to make your simulation get closer to the expectation?**

<div>
:::{#answer}

ANSWER HERE

:::
</div> 


---

##### Simulating the Independent Player

Simulating a basketball player who has independent shots uses the same mechanism that you used to simulate a coin flip. We will use `R` to make a *robot* basket ball player with the exact same shooting percentage as Kobe. Then we will ask the robot to shoot 133 baskets and compare the results to Kobe's results. We **know** that the robot **does not** have a hot hand, so Kobe should perform better than the robot if he does have a hot hand.

##### **Exercise 7. Make a vector called `shot_outcomes` that represent the possible outcomes from shooting a basketball.**

```{r}
#insert code here
```

##### **Exercise 8. Use the `sample()` function to simulate $133$ shots from a shooter with a shooting percentage of $45 \%$. Assign the output of this simulation to a new object called `sim_basket`.**

```{r}
#insert code here
```


---


##### **Exercise 9. Using `calc_streak()`, compute the streak lengths of `sim_basket`, and save the results in a data frame called `sim_streak`. **

```{r}
#insert code here
```




##### **Exercise 10. Describe the distribution of streak lengths. What is the average streak length for this simulated player with a $45 \%$ shooting percentage? How long is the player’s longest streak of baskets in $133$ shots? Make sure to include a plot in your answer. **

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>






##### **Exercise 11. If you were to run the simulation of the player a second time, how would you expect its streak distribution to compare to the distribution from the question above? Exactly the same? Somewhat similar? Totally different? Explain your reasoning.**

<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### **Exercise 12. This is just a single simulation and not really sufficient to compare Kobe's results to. Maybe the robot got lucky or unlucky in this simulation. We need more data. Replicate the process of finding the average streak length of $133$ shots from a $45 \%$ shooter $500,000$ times and save the result as `avg_steaks`. (This might take a little bit of time to run)**

<div>
:::{#answer}

ANSWER HERE

:::
</div>

The output of `replicate` is just a vector, but to do any graphing with `ggplot` we need a data frame. The code below will convert the atomic vector `avg_streaks` into a data frame with a column called `avg_streak`. Change `eval=FALSE` to `TRUE` and run the code:

```{r,eval=FALSE}
avg_streaks <- data.frame(avg_streak = avg_streaks)
```

##### **Exercise 13. (A) Plot a density histogram for the average streak length of the $500,000$ simulated shooters with a bin width of $.01$. (B) Describe the shape of the distribution.  (C) Add a blue vertical line for Kobe's average streak length and a red vertical line for the average of the average streak lengths. **

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### **Exercise 14. (A) Use `filter` to find all of the average shot streaks that are larger than Kobe's average shot streak. (B) What percentage of shooters have a larger average shot streak than Kobe?**

```{r}
#insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### **Exercise 15. What conclusions have you made about Kobe and the hot hand phenomenon?**

<div>
:::{#answer}

ANSWER HERE

:::
</div>

##### **Exercise 16. Do you think looking at the average streak length is the correct statistic to analyze? What could you change about the experiment in order to further investigate the hot hand phenomenon?**

<div>
:::{#answer}

ANSWER HERE

:::
</div>


