---
title: "Lab 2-Data Manipulation and Scatterplots"
author: Student 1 Name, Student 2 Name
output:
  html_document:
    css: ./style.css
    toc: no
    toc_float: no
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## Instructions

In this lab, you will practice manipulating data and creating scatterplots plots. 

### How to **complete** this assignment.

* Attempt each exercise in order.

* Please follow the Davidson Honor Code and rules from the course syllabus regarding seeking help with this assignment.

### How to **submit** this assignment.

* When you are finished, click the "Knit" button at the top of this panel and submit the HTML file in the same way you did for WS1. 

* This assignment is <font color="darkred">**due Friday, February 2, no later than 5:00 pm Eastern**</font>. Late submissions are not accepted without prior approval. 

---

#### **Exercise 1:**

+ **Load the `tidyverse` and `openintro` packages.** 
+ **Save the `nycflights` data set from the `openintro` package locally and name it `flights`.** 


```{r}
#insert code here
```


---

## Logical Operators

When you make logical statements with `R` it will respond with whether your statement is true or false. For example, if you ask `R` whether "to" is equal to "two" it will say false. Run the code below:

```{r}
"to" == "two"
```

Notice that we use two `=` sings for checking if things are equal (one equal sign is an assignment operator, similar to `->`).

Other operations `R` will check are:

+ greater than `>`
+ greater than or equal to `>=`
+ less than `<`
+ less than or equal to `<=`
+ equal to `==`
+ not equal to `!=`

You can also impose multiple conditions with the `&` (and) operator and `|` (or) operator. For example, you can ask `R` if $7$ is greater than or equal to $3$ with 

```{r}
7 >= 3
```

`R` outputs true. 

Now to use the `&` operator you can ask if $7$ is greater than $3$ **and** smaller than $5$:

```{r}
7 >= 3 & 7 < 5
```

We can also ask `R` if $7$ is greater than or equal to $3$ **or** $7$ is smaller than $5$
```{r}
7 >= 3 | 7 < 5
```

Now `R` outputs true.

#### **Exercise 2: Write `R` code that outputs True if $11^4$ is bigger than $8 \times 7 \times 6 \times 5 \times 4 \times 3 \times 2 \times 1$ and outputs false otherwise.**

```{r}
# insert code here
```

#### **Exercise 3: The code chunk below makes a vector of random numbers named `test_values`. Run the given code chunk then make another code chunk that will output a vector of `TRUE` or `FALSE` based on whether the numbers in `test_values` are bigger than $3$ and smaller than $20$**.

```{r}
test_values <- runif(15,1,100)
```

```{r}
# insert code here
```

---

## Filtering data

Often we will want to `filter()` a data set according to certain criteria. 

To use the `filter()` function you will pass in two arguments: 

+ the first argument is the data set 
+ the second argument is the condition you want to filter by. 

Let's `filter()` the `flights` data to find flights that departed and arrived on time:

```{r, eval=FALSE}
on_time_flights <- filter(flights, dep_delay == 0  | arr_delay == 0)
```

To find all flights that arrived on time but did not depart on time we would run the code:

```{r, eval=FALSE}
only_arrived_on_time <- filter(flights, dep_delay != 0  & arr_delay == 0)
```

To filter by anything by words instead of numbers you need to put quotes around the words: 

```{r, eval=FALSE}
filter(flights, origin == "JFK")
```

---

##### **Exercise 4. Find all flights that arrived more than $60$ minutes late. How many are there?**

```{r}
#insert code here
```

<div>
:::{#answer}

Answer here

:::
</div>

##### **Exercise 5. Locally save a new data matrix called `sfo_feb_flights` that only has flights headed to SFO in February. How many flights meet these criteria?** 

```{r}
#insert code here
```

<div>
:::{#answer}

Answer here

:::
</div>

##### **Exercise 6. What is the average arrival delay for `sfo_feb_flights`?**

```{r}
#insert code here
```

<div>
:::{#answer}

Answer here

:::
</div>

---

## Statistics by group

You can also calculate statistics for certain groups of your data. 

In order to get the average distance for all the flights we would use the command `mean(flights$distance)`.

However, in order to find the average distance **for each** month we need to `group_by()` month:

```{r, eval=FALSE}
summarize(group_by(flights,month),mean(distance,na.rm=TRUE))
```

---

##### **Exercise 7. Calculate the mean and median for arrival delays of flights in the `flights` data frame, grouped by carrier.**

+ **Which carrier has the highest mean arrival delay?**
+ **Which carrier has the highest median arrival delay?**

```{r}
#insert code here
```


##### **Exercise 8. Add a column to the data frame `flights` that has the average speed for each flight (in mph). Hint: Average speed can be calculated as distance divided by number of hours of travel, and note that `air_time` is given in minutes.**

```{r}
#insert code here
```



##### **Exercise 9. Make a scatterplot of average speed vs. distance.**  

```{r}
#insert code here
```


##### **Exercise 10. Add color to your scatter plot in order to distinguish between the origin airport.**

```{r}
#insert code here
```

##### **Exercise 11. Based on your scatterplot, does it seem that a particular airport flies shorter distance than the other two?**

<div>
:::{#answer}

Answer here

:::
</div>

### Penguins

##### **Exercise 12. Load the palmer penguin data and create a scatterplot that plots flipper length against body mass.**

```{r}
#insert code here
```

##### **Exercise 13. Add color to the graph based on the species of penguin.**

```{r}
#insert code here
```


##### **Exercise 14. What features of the graph from exercise 10 are more clear once color is added?**

<div>
:::{#answer}

Answer here

:::
</div>

##### **Exercise 15. Finally, split your plot into facets based on the island they live on?**

<div>
:::{#answer}

Answer here

:::
</div>

##### **Exercise 16. Which island has the fewest number of species on it?**

<div>
:::{#answer}

Answer here

:::
</div>

---
