---
title: "Worksheet 4 : Data Visualization"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(readr)
library(palmerpenguins)
library(openintro)
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
+ Scatter plots
+ Histograms
+ Skewness

---
  
## Scatter plots



1. Plot a scatter plot of the yrbss height vs. weight graph and add color to the graph according to grade.

```{r}
# insert code here
```



## Histograms

2. Make a vector called `height_in` by converting the heights to inches (there are $.0254$ meters in $1$ inch). 

```{r}
# insert code here
```

3. Add that column to the `teen_data`.

```{r}
# insert code here
```

4. Make a histogram for the heights of teens where the bin width is 1 inch.

```{r}
# insert code here
```

5. Try making a histogram for `race` in the yrbss data.

```{r}
# insert code here
```

### Density



## Skewness

Skew is a word we use to describe the shape of a distribution.

+ A **symmetric** distribution looks like:

```{r, echo = FALSE}
hist(rbeta(30000,5,5), main="Symmetric", xlab=NULL, ylab=NULL, xaxt='n', yaxt='n')
```

+ A **right** skew in a distribution looks like:

```{r, echo = FALSE}
hist(rbeta(30000,2,5), main="Right Skew", xlab=NULL, ylab=NULL, xaxt='n', yaxt='n')
```

+ A **left** skew in a distribution looks like:

```{r, echo = FALSE}
hist(rbeta(30000,5,2), main="Left Skew", xlab=NULL, ylab=NULL, xaxt='n', yaxt='n')
```


6. For each distribution shape (symmetric, right-skew, left-skew) compare the mean and median (eg. mean and median are about the same, mean is greater, median is greater)? 

<div>
:::{#answer}

Answer here

:::
</div>




When you have finished, Knit your project as an html document and submit the html document in Moodle with the following file name: "yourlastname_WS4.html"