---
title: "Worksheet 5 : Measures of Dispersion"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(palmerpenguins)
library(openintro)
library(knitr)
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Density Histograms
- Skewness
- Measures of Dispersion


---

## Density histograms


```{r}
teen_data <- yrbss
ggplot(teen_data, aes(x=height)) + geom_histogram(binwidth=.5, aes(y=after_stat(density)))
```

## Skewness

Skew is a word we use to describe the shape of a distribution.

+ A **symmetric** distribution looks like:

```{r, echo = FALSE}
df1 <- data.frame(x=rbeta(50000,5,5))
ggplot(df1,aes(x=x))+geom_histogram(binwidth = .025) + theme_void()
```

+ A **right** skew in a distribution looks like:

```{r, echo = FALSE}
df2 <- data.frame(x=rbeta(50000,2,5))
ggplot(df2,aes(x=x))+geom_histogram(binwidth = .025) + theme_void()
```

+ A **left** skew in a distribution looks like:

```{r, echo = FALSE}
df3 <- data.frame(x=rbeta(50000,5,2))
ggplot(df3,aes(x=x))+geom_histogram(binwidth = .025) + theme_void()
```


1. For each distribution shape (symmetric, right-skew, left-skew) compare the mean and median (eg. mean and median are about the same, mean is greater, median is greater)? Add a vertical line to the graphs to verify your answers.

<div>
:::{#answer}

Answer here

:::
</div>


## Measures of Dispersion


### Range


### Quantiles

### Interquartile Range 

### Box Plots



### Mean absolute deviation


$$ \bar{X}-X_i $$



$$| \bar{X}-X_i |$$


$$ \displaystyle \frac{\sum_{i=1}^n | \bar{X}-X_i |}{n}$$

Example: Consider the data set:

```{r}
example_data <- c(2, 4, 6, 28, 19, 29, 10)
```



So to compute the mean absolute deviation:

| data index $i$ | value $X_i$ | deviation from mean $\bar{X}-X_i$ | absolute deviation  $|\bar{X}-X_i|$|
| :---:|:---:|:---------:|:--------:|
| 1          |  2    |        ?            |          ?          |
| 2          |  4    |        ?            |          ?          |
| 3          |  6    |        ?            |          ?          |
| 4          |  28   |        ?            |          ?          |
| 5          |  19   |        ?            |          ?          |
| 6          |  29   |        ?            |          ?          |
| 7          |  10   |        ?            |          ?          |

2. Make a box plot for the flipper length data in palmer penguins.

```{r}
# insert code here
```

3. Estimate the IQR from the box plot then calculate the IQR exactly.

```{r}
#insert code here
```

4. Find the mean absolute deviation for the data set $\{ -6, 25, 77, 19, 144 \}$ without using the `aad()` function in the `lsr` package.

```{r}
#insert code here
```

5. Find the mean absolute deviation for the data set in exercise 4 using the `lsr` package.

```{r}
#insert code here
```

6. Find the median absolute deviation for the data set from exercise $4$.

```{r}
#insert code here
```

