---
title: "Worksheet 6 : Measures of Dispersion II"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
library(openintro)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Variance
- Samples and Populations
- Standard Deviation

---

Notes about lab:

+ make sure you are using ggplot for graphs
+ when you google how to change your graphs, make sure to include the word ggplot in your search
+ In the 17th problem it should say "add an appropriate title to your *box plots*"

## Review - Measures of Dispersion

So far, we have covered:

+ range, easy to calculate, not very descriptive
+ interquartile range, a little better than the range becuase it ignores outliers
+ mean and median absolute deviation (these are seldom used but do occassionally come up)
+ the most common measures of dispersion are the variance and standard deviation
  
## Variance


$$ Var(X) = s^2 =  \displaystyle \frac{\sum_{i=1}^n (\bar{X} - X_i)^2}{n} $$


---

1. Plot a histogram for the `yrbss` weight data. 

```{r}
# insert code here
```

2. Is the weight data symmetric, right-skewed, or left-skewed?

<div>
:::{#answer}

ANSWER HERE

:::
</div>

3. Find the mean, variance, and standard deviation for the weight data.

```{r}
# insert code here
```



4. What percentage of data lies within one standard deviation of the mean (find this exactly)?

```{r}
# insert code here
```

<div>
:::{#answer}

ANSWER HERE

:::
</div>

5. Is the percentage of data within one standard deviation of the mean more or less what you would expect using the general rule of thumb?

<div>
:::{#answer}

ANSWER HERE

:::
</div>

6. If a data set has mean $7$ and standard deviation $2$, how many standard deviations from the mean is the data point $11$? What about the data point $6$?

<div>
:::{#answer}

ANSWER HERE

:::
</div>