---
title: "Worksheet 7 : Standard Scores"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(knitr)
library(tidyverse)
library(openintro)
library(palmerpenguins)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Standardized Scores
- Probability


---


## Review - Standard Deviation

1. Load the `palmerpenguin` package.

```{r}
# insert code here
```

2. Estimate the population mean, population variance, and population standard deviation for the bill length of the penguins.

```{r}
# insert code here
```

3. Who performed better on their exam?

<div>
:::{#answer}
answer here
:::
</div>

  
## Standard Scores



## Probability

### Random Variables



4. What is the sample space is two different integers are chosen between $1$ and $5$ and listed in increasing order?

<div>
:::{#answer}
answer here
:::
</div>

5. What is the sample space if three dice are rolled and the sum is computed?

<div>
:::{#answer}
answer here
:::
</div>

