---
title: "Worksheet 1 : RStudio and Jupyter"
output:
  html_document:
    css: ./style.css
    toc: yes
    toc_float: yes
    theme: cosmo
  pdf_document:
    toc: yes
editor_options:
  chunk_output_type: console
---
  
```{r global_options, include = FALSE}
library(knitr)
knitr::opts_chunk$set(eval = TRUE, results = TRUE)
```

---
  
## Today's Agenda
- Exploring the RStudio Interface
- RMarkdown

---
  
During this worksheet we will explore R and RStudio. We’ll be using these throughout the course both to learn the statistical concepts discussed and to analyze real data and come to informed conclusions. 


+ R is a programming language. Allow us to give instructions to a computer.
+ RStudio is an interface, it makes coding in R easier.
+ A pro of programming: the computer can do tedious tasks quickly, and it will do exactly what we tell it to.
+ A con is that R will do exactly what you tell it to

## Exploring the RStudio Interface

+ Top Left: Workspace, take notes here. This is actually an RMarkdown file. This is like Markdown.
+ Bottom left: Console. where the magic happens. THis is base R
+ Top Right: Environment. Has saved data and variables.
+ Bottom Right: File and Package manager. It is also where plots will appear.

### R Console

Here is where R actually does stuff.

1. Does R follow PEMDAS (BEMDAS)? How can you test this in the console?

Answer here.

2. Type 3:7 into the console. What does the colon do in R?

Answer here.

3. Go to the console and press the up arrow key.What does this do?

Answer here.

### Workspace

Go to the top of the page and click knit

Adding R code to our notes, you can do this by clicking the green C at the top of the page and selecting the option R:

```{r}
3^5
```

Now press play and notice how this sends your R code down to the console to be executed.

Using R code chunks allows you to:
+ save and share work in a presentable way
+ Run a particular command again later
+ Make small edits without needing to retype everything.

4. Make an R code chunk and type 3:7 into it.

Answer here.

5. Make an R code chunk and evaluate $44(37)+129$. (Hint, if you type exactly what I put, you will get an error. Why is that happening, how can you fix it?)

Answer here.

